---
title: "Elections and Perceived Land Tenure in Zambia"
author: "Ben Ewing"
date: "January 24, 2019"
output: beamer_presentation
header-includes:
  - \usecolortheme[light]{solarized}
---

```{r packages, include = F}
library(dplyr)
library(tidyr)
library(lubridate)
library(readr)
library(purrr)
library(tidyr)
library(broom)
library(reggie)
library(glue)
library(sf)
library(ggplot2)
library(ggthemes)
theme_set(theme_minimal())
library(knitr)
opts_chunk$set(dev = 'pdf', echo = F)
```

```{r data, include = F}
tgcc <- read_csv("../data/private/processed/final.csv")
tgcc$chief_reallocate_any_fear <- tgcc$chief_reallocateh_bin
tgcc.wide <- tgcc %>% 
  gather(var, val, -c(hh_serial, time, vilid, treat, chief, matches("^t_"), matches("2015$"),
                      el_paperdoc_any, matches("2016$"))) %>% 
  unite(var, var, time) %>% 
  spread(var, val)

eastern_places <- readRDS("../data/private/interim/eastern_province_gis.rds")
eastern_border <- read_sf(
  "../data/raw/earthworks_zmb_consts_and_wards/Zambia_ConstituenciesAndWards_2006to2010_UTM35S.shp"
) %>% 
  filter(province == "Eastern")
eastern_border <- st_transform(eastern_border, crs = "+proj=longlat +datum=WGS84 +no_defs")
located_polls <- readRDS("../data/interim/polling_station_points.rds")
tgcc_polls <- located_polls %>% 
  filter(polling_station_url %in% c(tgcc$polling_station_url_2015, tgcc$polling_station_url_2016))
tgcc_polls_common <- tgcc_polls %>% 
  filter(polling_station_id %in% intersect(tgcc_polls$polling_station_id[tgcc_polls$year == 2015],
                                           tgcc_polls$polling_station_id[tgcc_polls$year == 2016]))
mean_sd <- function(x) {paste0("Mean: ", round(mean(x, na.rm = T), digits = 3), 
                               " SD: ", round(sd(x, na.rm = T), digits = 3))}
```

## Background

## Background

```{r background-timeline}
data_frame(
  date = ymd(c(
    "2014-06-01", "2014-08-01",
    "2017-06-16", "2017-08-09",
    "2014-10-28",
    "2015-01-20",
    "2016-08-11"
  )),
  event = c(
    "Start: Baseline Data Collection", "End: Baseline Data Collection",
    "Start: Midline Data Collection", "End: Midline Data Collection",
    "Death of Michael Sata",
    "2015 Election",
    "2016 Election"
  ),
  event_class = as.factor(c(
    "Baseline Data Collection", "Baseline Data Collection", 
    "Endline Data Collection", "Endline Data Collection",
    "Death of Michael Sata",
    "2015 Election",
    "2016 Election"
  ))
) %>% 
  ggplot(aes(y = 1, x = date, colour = event_class, label = event_class)) +
  geom_hline(yintercept = 1, size = 1.5) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  geom_text(check_overlap = T, vjust = 2) +
  xlim(ymd(c("2014-01-01", "2017-09-01"))) +
  scale_colour_few() +
  coord_fixed(2000) +
  theme_tufte() +
  theme(axis.text.y = element_blank(), axis.ticks = element_blank(),
        axis.title = element_blank(), legend.position="none") 
```

## Zambian Context

## Zambian Context -> Land

* Majority of land is under customary tenure
* At local level, land is administered by unelected Chiefs

## Zambian Context -> Elections -> 2015 and 2016

* 2015 Election
* Snap election after unexpected passing of president, Michael Sata
* No new voter registration $\Rightarrow$ used 2011 voter rolls
* 2016 Election
* Widespread voter registration $\Rightarrow$ updated voter rolls

## Zambian Context -> Elections -> Candidates

* Hichilema Hakainde for the United Party for National Development (UPND)
* **Edgar Lungu** for the Patriotic Front (PF)

## Zambian Context -> Eastern Province

* Where the data is from

## Data

## Data -> Election Data

* Polling station level vote tallies
* No exact polling station location

## Data -> GIS Data

* Compiled locations from many GIS datasets
* Matched location names to polling station names
* 2015 Election: `r length(unique(located_polls$polling_station_code[located_polls$year == 2015]))`
* Avg Turnout: `r mean(tgcc_polls$turnout_prop[tgcc_polls$year == 2015])`
* 2016 Election: `r length(unique(located_polls$polling_station_code[located_polls$year == 2016]))`
* Avg Turnout: `r mean(tgcc_polls$turnout_prop[tgcc_polls$year == 2016])`

## Data -> GIS Data -> Map

```{r map, cache = T, fig.height = 16, fig.width = 16, fig.align='center'}
ggplot() +
  geom_sf(data = eastern_border, colour = few_pal()(8)[3], fill = NA) +
  geom_sf(data = eastern_places, colour = few_pal()(8)[1], alpha = 0.075) +
  geom_sf(data = located_polls, colour = few_pal()(8)[4], alpha = 0.75) +
  geom_sf(data = tgcc_polls, colour = few_pal("Dark")(8)[6], alpha = 1) +
  theme(axis.text = element_blank())
```

## Data -> Survey Data

* TGCC RCT
* Baseline: 2014
* Endline: 2017
* N Panel Observations: `r nrow(tgcc.wide)`
* Tenure security, land investment, land governance indicators

## Data -> Survey Data -> Primary Outcomes

```{r tensec-plot}
tgcc %>% 
  select(any_fear, agroforest, fieldpurp_fallow, leader_fair_bin, leader_protect_bin, time) %>%
  rename(Year = time) %>% mutate(Year = factor(Year, c(0, 1), c(2014, 2017))) %>% 
  group_by(Year) %>% 
  summarise_all(funs(sum(., na.rm = T))) %>% 
  gather(var, val, -Year) %>% 
  ggplot(aes(x = var, y = val, fill = Year)) +
  scale_fill_few() +
  geom_col(position = "dodge") +
  ylab("N") +
  theme(axis.title.x =  element_blank())
```

## Data -> All Together Now

* Matched each village to nearest polling station
* 2016 Election: 
* N polling stations: `r length(unique(tgcc_polls$polling_station_code[tgcc_polls$year == 2016]))`
* N registered voters: `r as.character(sum(tgcc_polls$total_registered_voters[tgcc_polls$year == 2016]))`
* N votes: `r sum(tgcc_polls$total_votes_cast[tgcc_polls$year == 2016])`
* EL: `r mean_sd(tgcc_polls$lungu_edgar_c_PF[tgcc_polls$year == 2016]/tgcc_polls$total_votes_cast[tgcc_polls$year == 2016])`
* HH: `r mean_sd(tgcc_polls$hichilema_hakainde_UPND[tgcc_polls$year == 2016]/tgcc_polls$total_votes_cast[tgcc_polls$year == 2016])`

## Data -> All Together Now -> 2016 Election

```{r 2016-vote-dens}
tgcc_polls %>% 
  as.data.frame() %>% 
  select(lungu_edgar_c_PF, hichilema_hakainde_UPND) %>% 
  gather(Candidate, Votes) %>% 
  mutate(Candidate = factor(Candidate, levels = c("lungu_edgar_c_PF", "hichilema_hakainde_UPND"), 
                            labels = c("EL", "HH"))) %>% 
  ggplot(aes(Votes, colour = Candidate)) +
  geom_density() +
  scale_color_few()
```

## Results

## Results -> Model

Base model:

$y_{2016} = \beta_0 + \beta_1 y_{2015} + \beta_2 tAnyTenure + \beta_3 pollPropHH2016$

\noindent\rule{\textwidth}{1pt}

Paper document interaction model:

$y_{2016} = \beta_0 + \beta_1 y_{2015} + \beta_2 elPaperdocAny + \beta_3 pollPropHH2016 + \beta_4 (elPaperdocAny * pollPropHH2016)$

\noindent\rule{\textwidth}{1pt}

SEs clustered at polling station level

```{r model-functions}
build_models <- function(outcomes, fmla) {
  map_df(outcomes,
         function(y) {
           reg(tgcc.wide, glue(fmla),
               vcov_cluster = ~ polling_station_code_2016, vcov_type = "boot") %>% 
             summary() %>% 
             tidy() %>% 
             mutate(model = y)
         }) %>% 
    mutate(term = ifelse(grepl("_0$", term), "lagged", term),
           model = factor(model, levels = outcomes)) %>% 
    mutate(ci_low = estimate - std.error*qnorm(0.95), ci_high = estimate + std.error*qnorm(0.95))
}

model_plot_1 <- function(models, title, subtitle = "") {
  ggplot(models, aes(x = term, y = estimate, ymin = ci_low, ymax = ci_high)) + 
    geom_pointrange(colour = "#74a9cf") + 
    geom_hline(yintercept = 0, colour = "#0570b0", linetype = "dashed") + 
    facet_wrap( ~ model, ncol = 2) + 
    ggtitle(title) +
    coord_flip() + 
    theme_minimal() + 
    theme(axis.title = element_blank(), 
          panel.border = element_rect(colour = "#034e7b", fill = NA), 
          panel.spacing = unit(1, "lines"))
}

model_plot_2 <- function(models, title, subtitle = "") {
  models %>% 
    filter(term == "poll_prop_hh_2016") %>% 
    ggplot(aes(x = model, y = estimate, ymin = ci_low, ymax = ci_high)) + 
    geom_pointrange(colour = "#74a9cf") + 
    geom_hline(yintercept = 0, colour = "#0570b0", linetype = "dashed") + 
    ggtitle(title) + 
    coord_flip() + 
    theme_minimal() +
    theme(axis.title = element_blank(), 
          panel.border = element_rect(colour = "#034e7b", fill = NA), 
          panel.spacing = unit(1, "lines"))
}

model_plot_2_int <- function(models, title, subtitle = "") {
  models %>% 
    filter(term == "poll_prop_hh_2016" | term == "el_paperdoc_any" | term == "el_paperdoc_any:poll_prop_hh_2016") %>% 
    ggplot(aes(x = model, y = estimate, ymin = ci_low, ymax = ci_high)) + 
    geom_pointrange(colour = "#74a9cf") + 
    geom_hline(yintercept = 0, colour = "#0570b0", linetype = "dashed") +
    facet_grid(~ term) +
    ggtitle(title) + 
    coord_flip() + 
    theme_minimal() +
    theme(axis.title = element_blank(), 
          panel.border = element_rect(colour = "#034e7b", fill = NA), 
          panel.spacing = unit(1, "lines"))
}
```

```{r outcomes-and-models, cache = T, results = 'hide'}
tensec <- c(
  "any_fear", "elite_encroach_any_fear", "family_encroach_any_fear",
  "neighvill_encroach_any_fear", "villhh_encroach_any_fear", "headman_reallocate_any_fear",
  "chief_reallocate_any_fear"
)
investment <- c("agroforest", "fieldpurp_fallow")
landgov <- c("leader_fair_bin", "leader_protect_bin", "leader_transparent_bin", "leader_trust_bin", 
             "leaders_accountable_bin")

tensec_md <- build_models(tensec, "{y}_1 ~ {y}_0 + t_tenure + t_agro + t_agroXtenure + poll_prop_hh_2016")
tensec_interaction_md <- build_models(tensec, "{y}_1 ~ {y}_0 + t_any_agro + el_paperdoc_any + poll_prop_hh_2016 + el_paperdoc_any*poll_prop_hh_2016")
landgov_md <- build_models(landgov, "{y}_1 ~ {y}_0 + t_tenure + t_agro + t_agroXtenure + poll_prop_hh_2016")
investment_md <- build_models(investment, "{y}_1 ~ {y}_0 + t_tenure + t_agro + t_agroXtenure + poll_prop_hh_2016")
```

## Results -> Tenure Security

```{r tensec-results}
model_plot_2(tensec_md, "Effect of opposition party support on:", "Fear of encroachment indicators")
```

## Results -> Land Governance

```{r landgov-results}
model_plot_2(landgov_md, "Effect of opposition party support on:", "Land governance indicators")
```

## Results -> Land Investment

```{r investment-results}
model_plot_2(investment_md, "Effect of opposition party support on:", "Land investment indicators")
```

## Results -> Paper Documentation Interaction Models

```{r paper-doc-interaction}
model_plot_2_int(tensec_interaction_md, "Effect of opposition party support and land documentation on:")
```

## Future Work

* Better modeling approach
* Using 2015 and 2016 election data
* Within chiefdom models
* Locate more polling stations
